---
title: "Recreating 'Mean Field Games and Systemic Risk' by Carmona, Fouque, and Sun, 2013"
author: "Emma Holmes"
date: "March 2021"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r}
library(tidyverse)
set.seed(72021)
```

# Simple model
Here we model a system of N banks that borrow and lend to each other.
$$
dX^i_t = \frac{a}{N} \sum^N_{j=1} (X^j_t - X^i_t) dt + \sigma dW_t, \quad i=1,...,N
$$
where $W_t$ are independent Brownian motions and $X^i_0=0$ for $i=1,...,N$.

Here is code to simulate this simple model:
```{r}
###' @param N number of banks
###' @param a mean-reversion rate
###' @param sigma diffusion
###' @param X0 initial value for process X, assumed iid
###' @param end_time end of sim
###' @param dt Euler time step
simple_lending_model <- function(N,a,sigma,D=-0.7,rho=0,X0=0,end_time=1,dt=1e-4){
  ## set-up
  Nsteps <- end_time / dt + 1   # number of steps to take
  X <- matrix(nrow=N,ncol=Nsteps)   # empty matrix to store results
  X[,1] <- rep(X0,N)   # X_0 = X0 for all banks
  defaults <- rep(0,N)
  
  for (n in 2:Nsteps){
    dW0t <- sqrt(dt) * rnorm(1)   # common noise
    dWt <- sqrt(dt) * rnorm(N)   # Brownian motion
    X_bar <- sum(X[,n-1]) / N   # mean
    # Euler step:
    X[,n] = X[,n-1] + a * (rep(X_bar,N) - X[,n-1]) * dt + sigma *
      (rho * rep(dW0t,N) + sqrt(1-rho^2)*dWt)
    # record defaults
    defaults <- defaults + ifelse(X[,n]<= D,1,0)
  }
  # clean up defaults (1 if defaulted at least once)
  defaults <- ifelse(defaults>0,1,0)
  # clean up and return tibble of banks trajectories and times
  rownames(X) <- sapply(1:N, function(x) paste0("bank",x))
  times <- seq(0,end_time,by=dt)
  return(list(as_tibble(cbind(t(X),times)), defaults))
}
```

### Figure 2.1
Let $N=10$, and $\sigma=1$. Here are two sets of trajectories: one with $a=10$, and one with $a=0$, i.e. only Brownian motion. This recreates figure 2.1 from the paper.

```{r}
res1 = simple_lending_model(N=10,a=10,sigma=1,X0=0)
plot1 <- res1[[1]] %>% 
  pivot_longer(cols = 'bank1':'bank10',values_to = 'value',names_to = 'bank') %>% 
  ggplot(aes(times,value,colour=bank)) + geom_line(alpha=0.5) +
  geom_hline(yintercept = -0.7) + xlab('') + ylab('') +
  theme_bw() + theme(legend.position = "none") + ylim(-2,2) +
  ggtitle("a=10")

res2 = simple_lending_model(N=10,a=0,sigma=1,X0=0)
plot2 <- res2[[1]] %>% 
  pivot_longer(cols = 'bank1':'bank10',values_to = 'value',names_to = 'bank') %>% 
  ggplot(aes(times,value,colour=bank))+ geom_line(alpha=0.5) +
  geom_hline(yintercept = -0.7) + xlab('') + ylab('') +
  theme_bw() + theme(legend.position = "none") + ylim(-2,2) +
  ggtitle("a=0")

gridExtra::grid.arrange(plot1,plot2,nrow=1)
```

## Default distributions

This function computes the distribution of the number of defaults using the Monte Varlo method. 
```{r}
###' @param Nsims number of simulations for Monte Carlo
###' all other params as above
default_distribution <- function(Nsims,N,a,sigma=1,D=-0.7,rho=0,X0=0){
  # name file to save data:
  savefile <- sprintf("saved_sims/Nsims_%g_Nbanks_%g_a_%g_sigma_%g_rho_%g.Rdata",
                      Nsims,N,a,sigma,rho)
  if (!file.exists(savefile)){
      defaults <- rep(NA,Nsims)
      for (sim in 1:Nsims){
        # run the simple model Nsims times
        X <- simple_lending_model(N=N,a=a,sigma=sigma,D=D,rho=rho,X0=X0)
        # count number of defaults this run
        defaults[sim] <- sum(X[[2]])
      }
      save(defaults,file=savefile)
  } else{
      load(savefile)
    }
  return(defaults)
}
```

```{r}
Nsims <- 10000
defaults1 <- default_distribution(Nsims=Nsims,N=10,a=0)
defaults2 <- default_distribution(Nsims=Nsims,N=10,a=10)
defaults3 <- default_distribution(Nsims=Nsims,N=10,a=100)
```

### Figure 2.2
```{r, fig.height=4}
# figure 2.2 from paper
a1=as_tibble(defaults1) %>% count(value) %>% 
  mutate(n=n/Nsims) %>% rbind(c(value=10,n=0))
a2=as_tibble(defaults2) %>% count(value) %>% mutate(n=n/Nsims) 

fig_2_2a <- ggplot(data=a1,aes(value,n)) + 
  geom_line(colour='blue',lty=2) + 
  geom_line(data=a2, colour='blue') + 
  theme_classic() +
  xlab("# of default") + ylab("prob of # of default")

fig_2_2b <- ggplot(data=a1,aes(value,n)) + 
  geom_line(colour='blue',lty=2) + 
  geom_line(data=a2, colour='blue') + 
  theme_classic() + 
  coord_cartesian(xlim = c(6, 10),ylim=c(0,0.2)) +
  xlab("# of default") + ylab("prob of # of default")

gridExtra::grid.arrange(fig_2_2a,fig_2_2b,nrow=1)
```

```{r, fig.height=4}
# figure 2.2 with KDE instead
defaults12 <- rbind(mutate(as_tibble(defaults1),a = 0),
                  mutate(as_tibble(defaults2),a = 10)) 

plot_a <- defaults12 %>% mutate(a = as.factor(a)) %>%
  ggplot(aes(value, group=a,fill=a)) + 
  geom_density(adjust=2.5, alpha=0.5) + 
  theme_minimal() +
  xlab("# of default")

plot_b <- defaults12 %>% mutate(a = as.factor(a)) %>%
  ggplot(aes(value, group=a, fill=a)) + geom_density(adjust=2.5, alpha=0.5) + 
  theme_minimal() + coord_cartesian(xlim = c(6, 10),ylim=c(0,0.2)) +
  xlab("# of default")

gridExtra::grid.arrange(plot_a,plot_b,nrow=1)
```

### Figure 2.3
```{r, fig.height=4}
# figure 2.3 from paper
a3=as_tibble(defaults3) %>% count(value) %>% mutate(n=n/Nsims) 

fig_2_3a <- ggplot(data=a1,aes(value,n)) + 
  geom_line(colour='blue',lty=2) + 
  geom_line(data=a3, colour='blue') + 
  theme_classic() +
  xlab("# of default") + ylab("prob of # of default")

fig_2_3b <- ggplot(data=a1,aes(value,n)) + 
  geom_line(colour='blue',lty=2) + 
  geom_line(data=a3, colour='blue') + 
  theme_classic() + coord_cartesian(xlim = c(6, 10),ylim=c(0,0.2)) +
  xlab("# of default") + ylab("prob of # of default")

gridExtra::grid.arrange(fig_2_3a,fig_2_3b,nrow=1)
```

```{r, fig.height=4}
# figure 2.3 with KDE instead
defaults13 <- rbind(mutate(as_tibble(defaults1),a = 0),
                  mutate(as_tibble(defaults3),a = 100)) 

plot_c <- defaults13 %>% mutate(a = as.factor(a)) %>%
  ggplot(aes(value, group=a,fill=a)) + 
  geom_density(adjust=2.5, alpha=0.5) + 
  theme_minimal() +
  xlab("# of default")

plot_d <- defaults13 %>% mutate(a = as.factor(a)) %>%
  ggplot(aes(value, group=a, fill=a)) + geom_density(adjust=2.5, alpha=0.5) + 
  theme_minimal() + coord_cartesian(xlim = c(6, 10),ylim=c(0,0.2)) +
  xlab("# of default")

gridExtra::grid.arrange(plot_c,plot_d,nrow=1)
```

# Correlated Brownian Motions (figure 2.4)

```{r}
# Figure 2.4
res1 = simple_lending_model(N=10,a=10,sigma=1,rho=0,X0=0)
plot1 <- res1[[1]] %>% 
  pivot_longer(cols = 'bank1':'bank10',values_to = 'value',names_to = 'bank') %>% 
  ggplot(aes(times,value,colour=bank)) + geom_line(alpha=0.5) +
  geom_hline(yintercept = -0.7) + xlab('') + ylab('') +
  theme_bw() + theme(legend.position = "none") + ylim(-2,2) +
  ggtitle("rho=0")

res2 = simple_lending_model(N=10,a=10,sigma=1,rho=0.5,X0=0)
plot2 <- res2[[1]] %>% 
  pivot_longer(cols = 'bank1':'bank10',values_to = 'value',names_to = 'bank') %>% 
  ggplot(aes(times,value,colour=bank))+ geom_line(alpha=0.5) +
  geom_hline(yintercept = -0.7) + xlab('') + ylab('') +
  theme_bw() + theme(legend.position = "none") + ylim(-2,2) +
  ggtitle("rho=0.5")

gridExtra::grid.arrange(plot1,plot2,nrow=1)
```

## Systemic risk

Let $\{\min_{0\leq t \leq T} \overline X_t \leq D \}$ be the systemic event. Then 
$$ P \left( \min_{0\leq t \leq T} \overline X_t \leq D \right) = 2 \Phi\left( \frac{D\sqrt{N}}{\sigma \sqrt{T}} \right)$$

```{r}
N <- 10
D <- -0.7
end_time <- 1
sigma <- 1
prob_systemic <- 2*pnorm((D*sqrt(N))/(sigma * sqrt(T)))
```

So in the above simulations, the probability of the systemic event is `r round(prob_systemic,4)`, i.e. there is a 2.7% chance of the systemic event.


In the model with correlated BMs:

$$ P \left( \min_{0\leq t \leq T} \overline X_t \leq D \right) = 2 \Phi \left( \frac{D \sqrt{N}}{\sigma \sqrt{T(N-1) \rho^2 + 1}} \right)$$

the probability of the systemic event is 
```{r}
N <- 10
D <- -0.7
end_time <- 1
sigma <- 1
rho <- 0.5
prob_systemic2 <- 2*pnorm((D*sqrt(N))/(sigma * sqrt(T*(N-1)*rho^2 +1)))
```

The probability of the systemic event is now `r round(prob_systemic2,4)`

